Article Title

### 2013-10-05 sections/Estimators.tex

\subsubsection*{Estimator $\mathcal A$: Shrinkage toward diagonal.}
\emph{Estimator $\mathcal A$} shrinks the sample covariance matrix $\hat\Sigma_0$ toward that of the independent model shown in Fig.~\ref{fig:02}A.  The covariance matrix of a GGM in which all units are independent will be diagonal.  
For example, we  could leave the sample variances on the diagonal set the off-diagonal terms to zero. The target estimate
is then 
\begin{equation}\label{eq:sampleVariance}
\hat T= \hat\Sigma_0\circ I
\end{equation}
where $\circ$ denotes the entrywise matrix product (Hadamard product). 
Alternatively, the target could be a diagonal matrix with all variances equal to the average sample variance, $v = \frac 1 p \Tr(\hat\Sigma_0)$, so that
\begin{equation}\label{eq:equalVariance}
\hat T=\frac 1 p \Tr(\hat\Sigma_0) I
\end{equation}
The equal-variance target in Eq.~\ref{eq:equalVariance} has only one degree of freedom and thus lower estimation error but higher bias than the target with sample variances in Eq.~\ref{eq:sampleVariance}. It is not immediately clear which target is better in a particular application. We therefore use the linear mixture of the two targets paramaterized by the mixing coefficient $\alpha\in[0,1]$,
\begin{equation}
\hat T_\alpha = (1-\alpha)(\hat\Sigma_0 \circ I) + \frac \alpha p \mbox{tr}(\hat \Sigma_0)I.
\end{equation}
The estimator $\hat\Sigma_{\alpha,\lambda}$ is the linear shrinkage of the sample covariance matrix $\hat\Sigma_0$ toward $\hat T_\alpha$ controlled by the mixing proportion $\lambda\in[0,1]$,
\begin{equation}
\hat\Sigma_{\alpha,\lambda} = (1-\lambda)\hat\Sigma_0 + \lambda\hat T_\alpha.
\end{equation}
This covariance matrix estimator is implemented by R's {\tt corpcor} package \cite{Schaefer:2010} and is used widely~\cite{Schafer:2005}.

@@

As with estimator A, $\hat\Sigma_0$ is shrunk toward $\hat T_d$ by linear mixing controlled by shrinkage intensity $\lambda\in[0,1]$,
\begin{equation}
\hat\Sigma_{d,\lambda} = (1-\lambda)\hat\Sigma_0 + \lambda\hat T_d,
\end{equation}
and the optimal values of $d$ and $\lambda$ are found by cross-validation within the training set.

@@

The covariance matrix estimate is then 
\begin{equation}
-\hat\Sigma_\rho^{-1} =\arg\min\limits_S \loss{-S^{-1},\hat\Sigma_0}  \quad \mbox{such that} \quad  \|S_\rho\|_0 = \rho
\end{equation}
where $S_\rho$ is a sparse matrix with $\rho$ non-zero coefficients. $\|S\|_0$ denotes the number of non-zero elements of matrix $S$.
Solving this approximation in this form is computationally challenging. The problem can be made computationally tractable by replacing the $L_0$ norm with the $L_1$ norm, which converts the problem into one of convex optimization \cite{Donoho:2000}.  The resulting algorithm, known as \emph{graphical lasso} or \emph{GLASSO}  \cite{Meinshausen:2006,Yuan:2007,Banerjee:2008,Friedman:2008}, produces the following estimator  
\begin{equation}
-\hat\Sigma_\lambda^{-1} = \arg\min\limits_S \loss{-S^{-1},\hat\Sigma_0} + \lambda \|S\|_1
\end{equation}
where $\lambda \|S\|_1$ is the $L_1$ norm of the matrix. \TODO{clarify}. \Kcomment{Yes - at least define the norm, if you are already taking about it. No need to expand a lot more, though.}

@@

\subsubsection*{Estimator $\mathcal D$: Sparse inverse + low-rank}
\emph{Estimator $\mathcal D$} produces a low-dimensional estimate by finding an approximation of $\hat\Sigma_0$ approximating the sample covariance matrix such that its inverse is the sum of a spase matrix and a low-rank matrix. Covariance matrices of this form correspond to Gaussian graphical models of the form depicted in Fig.~\ref{fig:02}D.

The inverse $S=-\Sigma^{-1}$ of the covariance matrix of a GGM constitutes the graph of conditional dependencies (interactions) between the units.  With the assumption of sparse organization of interactions in the network, $S$ is assumed to be sparse.  When the observed units are part of a larger GGM that includes other latent units, let the negative inverse $S^*$ of the joint distribution of both observed and latent units be composed of four partitions
\begin{equation}
S^* = 
\begin{pmatrix}
S_{11} & S_{12} \\
S_{12}^\T & S_{22} 
\end{pmatrix}
\end{equation}
where partition $S_{11}$ describes the interactions between the visible units, $S_{12}$ describe the interactions between visible and latent units, and $S_{22}$ describes the interactions between the latent units. 
Then the covariance matrix of between the observed unit is 
\begin{equation}
\Sigma = -\left(S_{11} - S_{12}S_{22}^{-1}S_{12}^\T\right)^{-1}
\end{equation} 
The rank of matrix $S_{12}S_{22}^{-1}S_{12}$ equals the number of latent units. Rather than solving for the precise dependency structure with and between latent units in matrices $S_{12}$ and $S_{22}$, we simply replace the matrix $- S_{12}S_{22}^{-1}S_{12}$ with a low-rank matrix $L_d=- S_{12}S_{22}^{-1}S_{12}$ with rank $d$.

Then we obtain estimator $\hat\Sigma_{d,\rho}$, parameterized by the sparsity parameter $\rho$ and  the number of latent units $d$:
\begin{equation}
\begin{split}
-\hat\Sigma_{d,\rho}^{-1} &= \hat S_\rho + \hat L_d
\\
(\hat S_\rho, \hat L_d) &= \arg\min\limits_{S,L} \loss{-(S+L),\Sigma_0} \quad \mbox{such that} \quad \|S\|_0=\rho \quad \mbox{and} \quad \rank(L) = d
\end{split}
\end{equation}

Again, in this formulation, the problem is computationally intractible.  The problem can be made convex by replacing the $L_0$ norm with $L_1$ norm. The resulting algorith, known as \emph{latent-variable graphical lasso} or \emph{LV-GLASSO} \cite{Chandrasekaran:2010,Ma:2013}, produces estimator $\hat\Sigma_{d,\lambda}$:
\begin{equation}
\begin{split}
-\hat\Sigma_{d,\lambda}^{-1} &= \hat S_\lambda + \hat L_d
\\
(\hat S_\lambda, \hat L_d) &= \arg\min\limits_{S,L} \loss{-(S+L),\Sigma_0} + \lambda \| S \|_1 \quad \mbox{such that}  \quad \rank(L) = d
\end{split}
\end{equation}
The optimal values of hyperparameters $\lambda$ and $d$ are found cross-validation within the training dataset.

>>>>

\subsubsection*{Estimator $\mathcal A$: Shrinkage toward diagonal.}
\emph{Estimator $\mathcal A$} shrinks the sample covariance matrix $\hat\Sigma_0$ toward that of the independent model shown in Fig.~\ref{fig:02}A.  The covariance matrix of a GGM in which all units are independent will be diagonal.  
For example, we  could leave the sample variances on the diagonal set the off-diagonal terms to zero. The target estimate
is then 
\begin{equation}\label{eq:sampleVariance}
\hat T= \hat\Sigma_0\circ I
\end{equation}
where $\circ$ denotes the entrywise matrix product (Hadamard product). 
Alternatively, the target could be a diagonal matrix with all variances equal to the average sample variance, $v = \frac 1 p \Tr(\hat\Sigma_0)$, so that
\begin{equation}\label{eq:equalVariance}
\hat T=\frac 1 p \Tr(\hat\Sigma_0) I
\end{equation}
The equal-variance target in Eq.~\ref{eq:equalVariance} has only one degree of freedom and thus lower estimation error but higher bias than the target with sample variances in Eq.~\ref{eq:sampleVariance}. It is not immediately clear which target is better in a particular application. We therefore use the linear mixture of the two targets paramaterized by the mixing coefficient $\alpha\in[0,1]$,
\begin{equation}
\hat T_\alpha = (1-\alpha)(\hat\Sigma_0 \circ I) + \frac \alpha p \mbox{tr}(\hat \Sigma_0)I.
\end{equation}
The estimator $\hat\Sigma_{\alpha,\lambda}$ is the linear shrinkage of the sample covariance matrix $\hat\Sigma_0$ toward $\hat T_\alpha$ controlled by the mixing proportion $\lambda\in[0,1]$,
\begin{equation}
\hat\Sigma_{\alpha,\lambda}^\mathcal{A} = (1-\lambda)\hat\Sigma_0 + \lambda\hat T_\alpha.
\end{equation}
This covariance matrix estimator is implemented by R's {\tt corpcor} package \cite{Schaefer:2010} and is used widely~\cite{Schafer:2005}.

@@

Just as with estimator $\mathcal A$, $\hat\Sigma_0$ is shrunk toward $\hat T_d$ by linear mixing controlled by shrinkage intensity $\lambda\in[0,1]$,
\begin{equation}
\hat\Sigma_{d,\lambda}^\mathcal{B} = (1-\lambda)\hat\Sigma_0 + \lambda\hat T_d,
\end{equation}
and the optimal values of $d$ and $\lambda$ are found by cross-validation within the training set.

@@

The covariance matrix estimate is then 
\begin{equation}
\hat\Sigma_\rho = \hat S^{-1}
\quad\mbox{with}\quad
\hat S = \argmin\limits_{\substack{S\succ 0 \\ \|S\|_0 \le \rho}} \loss{S^{-1},\hat\Sigma_0}   
\end{equation}
where $S\succ 0$ signifies the constraint that $S$ be positive-definite. $\|S\|_0\le\rho$ signifies the constraint  that matrix $S$ have at most $\rho$ non-zero coefficients.
Solving this approximation in this form is computationally challenging. The problem can be made computationally tractable by relaxing \TODO{use "complex relaxation"} the $L_0$ norm to the $L_1$ norm, which converts the problem into one of convex optimization \cite{Donoho:2000}.  The resulting algorithm, known as \emph{graphical lasso} or \emph{GLASSO}  \cite{Meinshausen:2006,Yuan:2007,Banerjee:2008,Friedman:2008}, produces the following estimator  
\begin{equation}
\hat\Sigma_\lambda^\mathcal{C} = \hat S^{-1}
\quad\mbox{with}\quad
\hat S = \argmin\limits_{S\succ 0} \loss{S^{-1},\hat\Sigma_0} + \lambda \|S\|_1
\end{equation}
where $\lambda \|S\|_1$ is the $L_1$ norm of the matrix. \TODO{clarify}. \Kcomment{Yes - at least define the norm, if you are already taking about it. No need to expand a lot more, though.}

@@

\subsubsection*{Estimator $\mathcal D$: Sparse + low-rank inverse}
\emph{Estimator $\mathcal D$} produces a low-dimensional estimate by approximating $\hat\Sigma_0$ with a matrix whose inverse is the sum of a spase matrix and a low-rank matrix. Covariance matrices of this form correspond to Gaussian graphical models of the form depicted in Fig.~\ref{fig:02}D.

The inverse $S=\Sigma^{-1}$ of the covariance matrix of a GGM constitutes the graph of conditional dependencies (interactions) between the units.  With the assumption of sparse organization of interactions in the network, $S$ is assumed to be sparse.  When the observed units are part of a larger GGM that includes other latent units, let the negative inverse $S^*$ of the joint distribution of both observed and latent units be composed of four partitions
\begin{equation}
S^* = 
\begin{pmatrix}
S_{11} & S_{12} \\
S_{12}^\T & S_{22} 
\end{pmatrix}
\end{equation}
where partition $S_{11}$ describes the interactions between the visible units, $S_{12}$ describe the interactions between visible and latent units, and $S_{22}$ describes the interactions between the latent units. 
Then the covariance matrix of between the observed unit is 
\begin{equation}
\Sigma = \left(S_{11} - S_{12}S_{22}^{-1}S_{12}^\T\right)^{-1}
\end{equation} 
The rank of matrix $S_{12}S_{22}^{-1}S_{12}$ equals the number of latent units. Rather than solving for the precise dependency structure with and between latent units in matrices $S_{12}$ and $S_{22}$, we simply replace the matrix $ S_{12}S_{22}^{-1}S_{12}$ with a low-rank matrix $L= S_{12}S_{22}^{-1}S_{12}$.

Then we obtain estimator $\hat\Sigma_{d,\rho}$, parameterized by the sparsity parameter $\rho$ and  the number of latent units $d$:
\begin{equation}
\hat\Sigma_{d,\rho} = \left(\hat S-\hat L\right)^{-1}
\quad\mbox{where}\quad
(\hat S, \hat L) = \argmin\limits_{\substack{S,L\\ (S-L)\succ 0 \\ \|S\|_0\le\rho \\ L\succeq 0 \\ \rank(L) \le d}} \loss{(S-L)^{-1},\Sigma_0}
\end{equation}

Again, in this formulation, the approximation problem is computationally intractible.  The problem can be made convex by replacing the $L_0$ norm with $L_1$ norm. The resulting algorith, known as \emph{latent-variable graphical lasso} or \emph{LV-GLASSO} \cite{Chandrasekaran:2010,Ma:2013}, produces estimator $\hat\Sigma_{d,\lambda}^{\mathcal D}$:
\begin{equation}
\hat\Sigma_{d,\lambda} ^{\mathcal D} = \left(\hat S-\hat L\right)^{-1}
\quad\mbox{where}\quad
(\hat S, \hat L) = \argmin\limits_{\substack{ S,L \\ (S-L)\succ 0 \\ L\succeq 0 \\ \rank(L) \le d}} \loss{(S-L)^{-1},\Sigma_0} + \lambda \| S \|_1 
\end{equation}
The optimal values of hyperparameters $\lambda$ and $d$ are found cross-validation within the training dataset.

### 2013-10-05 sections/Figure2.tex

\begin{figure}[htp]
\centering
\includegraphics[width=0.5\textwidth]{figures/Figure2.pdf}
\caption{
Graphical models corresponding to the low-dimensional targets of the four regularization schemes used in the paper.
\textbf{A}: A diagonal matrix corresponds to a Gaussian graphical model with no dependencies.
\textbf{B}: In factor analysis, observed nodes are assumed to be influenced by several latent units (``factors") but are otherwise independent.
\textbf{C}: In the Gaussian graphical model (also known as the Gaussian Markov Field), correlations arise from sparse pairwise linear interactions between visible units.
\textbf{D}: In the Gaussian graphical model with latent units, correlations arise  between pairs of nodes
}\label{fig:02}
\end{figure}

>>>>

\begin{figure}[htp]
\centering
\includegraphics[width=0.5\textwidth]{figures/Figure2.pdf}
\caption{
Graphical models corresponding to the low-dimensional targets of the four regularization schemes used in the paper.
\textbf{A}: A diagonal matrix corresponds to a Gaussian graphical model with no dependencies.
\textbf{B}: In factor analysis, observed nodes are assumed to be influenced by several latent units (``factors") but are otherwise independent.
\textbf{C}: Graphical model with conditional dependencies between some pairs of observed neurons, no latent units. 
\textbf{D}: Graphical model with conditional dependencies between some pairs of observed neurons and dense interactions with a few latent units.
}\label{fig:02}
\end{figure}

### 2013-10-05 sections/Introduction.tex
Added Author Summary

With large numbers of recorded cells, the usual estimations of the covariance matrix from finite recordings become ill-conditioned. Numerical instabilities arise because, as the amount of recorded data increases only linearly, the number of free coefficients to be estimated in any parameterization of the covariance matrix increases quadratically. 
For example, the large eigenvalues of large sample covariance matrices are biased upward while small eigenvalues are biased downward \cite{Hayes:1981}. Similarly, the coefficients of the inverse covariance matrix require larger sample sizes to be estimated accurately from high-dimensional data than low-dimensional data.

With this study, we pursue two related goals: 
\begin{enumerate}
	\item Devise more efficient estimators of neural covariance matrices in recordings from large populations of neurons.
	\item Aid interpretation of population activity  by isolating the essential structure of the neural covariance matrix.
\end{enumerate}

>>>>

With large numbers of recorded cells, the usual estimations of the covariance matrix from finite recordings become ill-conditioned. Numerical instabilities arise because, as the amount of recorded data increases only linearly, the number of free coefficients to be estimated in any parameterization of the covariance matrix increases quadratically. 
For example, the large eigenvalues of large sample covariance matrices are biased upward while small eigenvalues are biased downward \cite{Ledoit:2004}. Similarly, the coefficients of the inverse covariance matrix require larger sample sizes to be estimated accurately from high-dimensional data than low-dimensional data.

With this study, we pursue two related goals: 
\begin{enumerate}
	\item Devise more efficient estimators of neural covariance matrices in recordings from large populations of neurons.
	\item Aid interpretation of neural covariance matrices.
\end{enumerate}
